---
title: "Skill Mismatch & Labour Market Outcomes"
author: "Odud Bin Yousuf"
date: "`r Sys.Date()`"
---

```{r, echo=TRUE}
rm(list=ls()); cat('\014')
```

Installing the relevant packages and importing the relevant libraries. 
```{r,echo=TRUE}

#Undo the hashtages below if you need to install anyone of the packages below. 
#      I have already installed them so have the hashtag 

#install.packages("haven")
#install.packages("rio")
#install.packages("plm")
#install.packages("lmtest")
#install.packages("sandwich")
#install.packages("stargazer")
#install.packages("xtable")
#install.packages("knitr")
#install.packages("psych")
#install.packages("kableExtra")
#install.packages("flextable")
#install.packages("moments")
#install.packages("polycor")

#Importing the relevant libraries
library(haven)
library(rio)
library(readxl)
library(dplyr)
library(purrr)
library(plm)
library(lmtest)
library(sandwich)
library(stargazer)
library(xtable)
library(knitr)
library(psych)
library(ggplot2)
library(gridExtra)
library(moments)
library(tseries)
library(polycor)
```

Setting the Working directory.(set your own directory)
```{r setup,echo=TRUE,}
#Setting the working directory 
getwd()
#setwd(add working directory here)
```

Reading the relevant datasets of 'Work and Schooling' from 2019 to 2022 and the background variables (of 2022).  
```{r,echo=TRUE}

#Reading datasets from 2019 to 2022
data2022 <- haven::read_dta("cw22o_EN_1.0p.dta")
data2021 <- haven::read_dta("cw21n_EN_1.0p.dta")
data2020 <- haven::read_dta("cw20m_EN_1.0p.dta")
data2019 <- haven::read_dta("cw19l_EN_3.0p.dta")  
background2022<- haven::read_dta("avars_202204_EN_1.0p.dta")

```


Filtering the dataset and removing unwanted variables. Removing the prefixes from the variable names so they have the same names for each dataset. 
```{r,echo=TRUE}

#Deleting unused/irrelevant variables from dataset: 
##Deleting from 2022

to_delete_2022<- c(002,008,009,029,035,037,040,042:050, 052,053, 105:110,112,534,
                   114:118,511,535,536,120,319:321, 121,537,538,124,125,122,123,131, 134:136,610,
                   146:222, 234:286, 288,289:306, 308:311, 313:316, 318,322,340:360, 362:363,611,
                   405:407, 528,409:411, 426:435, 441:443, 529, 450:454,557:570,572,578,579,
                   581:587,588:597, 599:602, 604:609, 496:500, 501:505)

variables_to_delete_2022 <- paste0("cw22o", sprintf("%03d", to_delete_2022 ))


data2022_clean <- data2022[, !(names(data2022) %in% variables_to_delete_2022)]

varsleft<-c(print(names(data2022_clean)))
varsleft_clean <- gsub("cw22o", "", varsleft)


#Renaming the variables to remove the "cwyym" where yy stands for the year 
#                                                             of the dataset and m for month


new_names19 <- gsub("^cw19l", "", names(data2019))
names(data2019) <- new_names19
new_names20 <- gsub("^cw20m", "", names(data2020))
names(data2020) <- new_names20
new_names21 <- gsub("^cw21n", "", names(data2021))
names(data2021) <- new_names21
new_names22 <- gsub("^cw22o", "", names(data2022_clean))
names(data2022_clean) <- new_names22


#Deleting/ only keeping the relevant variables from all the datasets 
data2021_clean <- data2021[, varsleft_clean ]
data2020_clean <- data2020[, varsleft_clean ]

#005 variable is not in 2019 dataset:
data2019_clean <- data2019[, varsleft_clean[varsleft_clean!="005"] ] 

```


Combining datasets to make the panel datasets. 
```{r,echo=TRUE}
#Combining the relevant data into 1 panel data:

suppressWarnings({
  panel_data21_22 <- list(data2021_clean,data2022_clean) %>% reduce(full_join, by = varsleft_clean)
})

suppressWarnings({
panel_data20to22 <- list(data2020_clean,panel_data21_22) %>% reduce(full_join, by = varsleft_clean)
})


suppressWarnings({
  panel_data19to22 <- list(data2019_clean,panel_data20to22) %>% reduce(full_join, by = varsleft_clean[varsleft_clean!="005"])
}) #this line takes a while to run 
```

Clear my global environment from stuff I don't need. 
```{r,echo=TRUE}
#remove datasets and variables I don't need now since they are inside panel_data19to22:
rm(data2019_clean,data2020_clean,data2021_clean,panel_data21_22,new_names19,new_names20 
   ,new_names21, new_names22)
#removing unclean datasets from global environment:
rm(panel_data20to22,data2022,data2021,data2020,data2019, data2022_clean)
```


Reducing to get the final 1 panel dataset that I will use for all my data analysis.
```{r,echo=TRUE}
#REDUCING THE PANEL DATA INTO THE RELEVANT one and add gender 

#a]
#Removing all working students from my dataset: 
##counting working students: 
sum(panel_data19to22$'523' %in% 1) 
sum(panel_data19to22$'523' %in% 1) /22109
# = 494 out of 22109 are work students. So 2.23% of dataset are work students 

#b]
#Removing all respondent over the age of 66 as that is the retirement age in NL:
#over retirement age:
sum(panel_data19to22$'003' > 66, na.rm = TRUE ) 
sum(panel_data19to22$'003' > 66 ) /22109
# = 6171 out of 22109 are over 66. 
#over retirement age and unemployed:
sum(panel_data19to22$'003' > 66 & panel_data19to22$'000' %in% 0) 
sum(panel_data19to22$'003' > 66 & panel_data19to22$'000' %in% 0) / 22109
# Even though 27.91171% are over 66, 27.10061% of that are unemployed. 

panel_data19to22 <- panel_data19to22[panel_data19to22$'523' != 1, ] #1st remove work student
panel_data19to22 <- panel_data19to22[panel_data19to22$'003' <= 66, ] #then retirees

#Reducing our dataset from 22109 to 15447 

#c]
#Adding the gender 
paneldt <- merge(panel_data19to22, background2022[, c("nomem_encr", "geslacht")]
                     , by = "nomem_encr", all.x = TRUE)

#d]Keeping the needed variables in the dataset and removing the irrelevant ones

final_variables <- c('nomem_encr','_m','006','577','001','127','003','402','005','031', '549'
                     ,'550', '551', '552', '553', '554', '555', '033','128','129','145','143','522','144', 'geslacht')


paneldt<- paneldt[, final_variables]

```

I remove all observations where the worker is unemployed or long term unemployed since these observations do not add any value to my analysis and are therefore irrelevant. These observations also produce too many missing values. I also drop those with a side-job for similar reasons. 
```{r,echo=TRUE}
sum(paneldt$`001` %in% 0 ) # = 4122 observations (26.79% of my datset rn is unemployed)
#removing unemployed & long term unemployed respectively. 
paneldt <- paneldt[paneldt$'001' != 0, ]
paneldt <- paneldt[paneldt$'522' != 1, ]

##########Dropping those with a side job 
paneldt <- paneldt[!(paneldt$'143' %in% c(1, 2)), ]
```


Dealing with missing values: 
```{r,echo=TRUE}

#Those missing values for observations that have a job but NA for side-job variables 

sum(paneldt$`143` %in% NA) # 237  NAs 
# - Turn the NAs that correspond to them having a job into 3 (no side job) 
#                                since they do have a job but have NA for sidejob. 
sum(paneldt$`143` %in% NA & paneldt$'001' %in% 1)
paneldt$'143'[paneldt$`001` == 1 & is.na(paneldt$'143')] <- 3


#- remove the rest 58 NAs that are due to merely no response 
paneldt <- paneldt[complete.cases(paneldt$'143'),]
#!!!!!dataset is reduced to 10392 (NOTE:later reduced to 10131 when dealing w NAs of wages & hours)

# I can also now remove the variable '143' and '144'
#             from the dataset since I don't need it anymore 
paneldt <- paneldt[, !(names(paneldt) == "143")]
paneldt <- paneldt[, !(names(paneldt) == "144")] #don't need 144 either 


```


Problems with the independent variables of interest (educational and skill mismatch variables)
```{r,echo=TRUE}
#a] Skill mismatch variables: Making the "I don't know" responses into "my skills are insufficiently geared for my work" i.e. differently skilled 


#Checking -9s in 549 to 555 = skill matches & 033 
sum(paneldt$'549' %in% -9)
sum(paneldt$'550' %in% -9)
sum(paneldt$'551' %in% -9)
sum(paneldt$'552' %in% -9)
sum(paneldt$'553' %in% -9)
sum(paneldt$'554' %in% -9)
sum(paneldt$'555' %in% -9)
# = 133 for each 
#Are the same respondents putting idk for all the 7 variables? 
paneldt$'nomem_encr'[paneldt$'554' == -9]
identical(paneldt$'nomem_encr'[paneldt$'549' == -9], paneldt$'nomem_encr'[paneldt$'555' == -9])
#Yes! The same respondents are answering idk for 549 to 555
sum(paneldt$`033` %in% 99) # = 4 idks for those who answered twice or more for 549-555 (='033')

#Change the idks to 1s in 555:
paneldt$'549'[paneldt$'549' == -9] <- 0
paneldt$'550'[paneldt$'550' == -9] <- 0
paneldt$'551'[paneldt$'551' == -9] <- 0
paneldt$'552'[paneldt$'552' == -9] <- 0
paneldt$'553'[paneldt$'553' == -9] <- 0
paneldt$'554'[paneldt$'554' == -9] <- 0
paneldt$'555'[paneldt$'555' == -9] <- 1

#b] Educational Mismatch variables: Making the "I don't know" responses into "my education level is insufficiently geared for my work" i.e. differently educated 

sum(paneldt$`031` %in% 99) # = 199
paneldt$'031'[paneldt$'031' == 99] <- 7

```



Defining the variables as I will use them in my thesis study: 
```{r,echo=TRUE}
#Defining the educ variables based on 031=my education... 
paneldt$edumatch <- ifelse(paneldt$'031' %in% 1 , 1 , 0)
paneldt$overedu  <- ifelse(paneldt$'031' %in% 2 , 1 , 0)
paneldt$underedu <- ifelse(paneldt$'031' %in% 3 , 1 , 0)
paneldt$otheredu <- ifelse(paneldt$'031' %in% c(4,5,6,7 ), 1 , 0)


#Defining the skill variables 
paneldt$skillmatch   <- ifelse(paneldt$'549' %in% 1, 1, 0)
paneldt$overskilled  <- ifelse(paneldt$`550` %in% 1, 1, 0)
paneldt$underskilled <- ifelse(paneldt$`551` %in% 1, 1, 0)
paneldt$otherskill   <- ifelse(paneldt$'552' == 1 | paneldt$'553' == 1 |paneldt$'554' == 1 | paneldt$'555' == 1, 1, 0)  


#Renaming '003' age and 'geslacht' gender
paneldt <- paneldt %>% rename(age = '003' )
paneldt <- paneldt %>% rename(gender = 'geslacht' )

#Defining the male and female variables 
paneldt$male     <- ifelse(paneldt$gender %in% 1, 1, 0)
paneldt$female   <- ifelse(paneldt$gender %in% 2, 1, 0)

#Defining the sectors that the respondent works in based on 402 
paneldt$primary  <- ifelse(paneldt$'402' %in% c( 1, 2), 1, 0)
paneldt$secondary <- ifelse(paneldt$'402' %in% c(3, 4, 5), 1, 0)
paneldt$tertiary1 <- ifelse(paneldt$'402' %in% c(6,7,8), 1, 0)
paneldt$tertiarycommerce <- ifelse(paneldt$'402' %in% c( 9, 10), 1, 0)
paneldt$tertiary3 <- ifelse(paneldt$'402' %in% c( 11,12,13,14), 1, 0)
#Rename 402 sector 
paneldt <- paneldt %>% rename(sector = '402' )

#Defining the highest level of education achieved using 005 & 006
paneldt$educlevel_barelyany <- ifelse(paneldt$'005' %in% c(1, 2) | paneldt$'006' %in% 1, 1, 0)
paneldt$educlevel_prehigh <- ifelse(paneldt$'005' %in% c(3, 4, 5, 6, 7) | paneldt$'006' %in% c(2, 3), 1, 0)
paneldt$educlevel_high    <- ifelse(paneldt$'005' %in% c(8, 9, 10, 11, 12, 13, 14) | paneldt$'006' %in% 4, 1, 0)
paneldt$educlevel_further <- ifelse(paneldt$'005' %in% c(15, 16, 17, 18, 19, 20, 21, 22) | paneldt$'006' %in% 5, 1, 0)
paneldt$educlevel_bachelor <- ifelse(paneldt$'005' %in% c(23,25) | paneldt$'006' %in% 6, 1, 0)
paneldt$educlevel_masters <- ifelse(paneldt$'005' %in% c(24, 26) | paneldt$'006' %in% 7, 1, 0)
paneldt$educlevel_phd     <- ifelse(paneldt$'005' %in% 27 , 1, 0)


#Renaming hours, wage for 127 and 577 respectively. 
paneldt <- paneldt %>% rename(hours = '127' )
paneldt <- paneldt %>% rename(wage = '577' )
#Renaming nomem_encr and _m as resp_id and date respectively  
paneldt <- paneldt %>% rename(resp_id = nomem_encr )
paneldt <- paneldt %>% rename(date = "_m" )

#FINAL REMOVAL of the variables I don't need now since I have already used them in defining the variables I will use
to_delete_final <- c('006', '005', '031', '549', '550', '551', '552', '553', '554', '555'
                     , '033', '128', '129', '145', '001', '522', 'gender')
paneldt <- paneldt[, !(names(paneldt) %in% to_delete_final)]
```


Dealing with the missing and problem values in our dependent variables: wage and hours
```{r,echo=TRUE}

#How many NAs & idks in hours = avg HRS: 
sum(paneldt$`hours` %in% NA)    # =111 (= 1.1% of sample)
sum(paneldt$'hours' %in% 999)  # = 0 idks in hours 

#How many NAs, idks & prefer not to say (pns) in 577 = gross monthly wage
sum(paneldt$`wage` %in% NA) # =2324  NAs
sum(paneldt$`wage` %in% -9) # =1176 idks
sum(paneldt$`wage` %in% -8) # =1599 pns
#around 50% of the dataset would be removed if i removed these NAs,idks & pns for wage. 

#Removing NAs, -9s and -8s for wage
paneldt <- paneldt[!(paneldt$wage %in% c(NA, -9, -8)), ]
#Removing NAs for hours 
paneldt <- paneldt[!(paneldt$hours %in% NA), ]


#Only 9 missing values left in the data which are removed 
paneldt <- na.omit(paneldt)


#Remove sector from dataset as I don't need it anymore
paneldt <- paneldt[, -which(names(paneldt) == "sector")]

#Remove the  irrelevant stuff from my global environment
rm(background2022)
```



Defining logwage
```{r,echo=TRUE}
paneldt$logwage <- log(paneldt$wage)
```

Making my final dataset into a panel dataset
```{r,echo=TRUE}
pdt <- pdata.frame(paneldt , index = c("resp_id" , "date"))
```
This is our final sample panel dataset that will be used for all regressions and data analysis. The indexes are the respondent ID and date which correspond to the i and t in regressions since this is how the data should be. We have a final N=5,277 observations. 


Summary statisitcs : 
```{r,echo=TRUE}
sum1<- describe(pdt[, c('wage', 'hours','age', 'female','male', 'edumatch', 'overedu', 'underedu', 'otheredu',
                 'educlevel_prehigh', 'overskilled', 'underskilled', 'otherskill', 'skillmatch',
'educlevel_high', 'educlevel_further', 'educlevel_bachelor', 'educlevel_masters', 'educlevel_phd'
      ,'primary', 'secondary', 'tertiary1', 'tertiarycommerce', 'tertiary3') ])


print(sum1[, c("mean", "sd", "n")])
```



REGRESSIONS: 

1] Pooled OLS regressions 

```{r,echo=TRUE}
#pooled ols with hours 
pooled_1 <- plm (hours ~ logwage + age + female + educlevel_prehigh +educlevel_high + educlevel_further
                         + educlevel_bachelor + educlevel_masters + educlevel_phd
                         + primary + secondary + tertiary1 +tertiarycommerce
                         + overedu + underedu + otheredu 
                         + overskilled + underskilled + otherskill , data = pdt, model = "pooling" )

summary(pooled_1)
#pooled ols with logwages
pooled_2 <- plm (logwage ~ hours + age + female + educlevel_prehigh +educlevel_high + educlevel_further
                 + educlevel_bachelor + educlevel_masters + educlevel_phd
                 + primary + secondary + tertiary1 +tertiarycommerce
                 + overedu + underedu + otheredu 
                 + overskilled + underskilled + otherskill , data = pdt, model = "pooling" )

summary(pooled_2)
```

2]Fixed Effect regresssions
```{r,echo=TRUE}
femethod_1 <- plm( hours ~ logwage 
                 + overedu + underedu + otheredu 
                 + overskilled + underskilled + otherskill , data = pdt, model = "within" )

summary(femethod_1)

femethod_2 <- plm(logwage ~ hours 
                  + overedu + underedu + otheredu 
                  + overskilled + underskilled + otherskill , data = pdt, model = "within" )

summary(femethod_2)
```

3]Random Effect regressions
```{r,echo=TRUE}
remethod_1 <- plm(hours ~ logwage + age + female + educlevel_prehigh +educlevel_high + educlevel_further
                + educlevel_bachelor + educlevel_masters + educlevel_phd
                + primary + secondary + tertiary1 +tertiarycommerce
                + overedu + underedu + otheredu 
                + overskilled + underskilled + otherskill , data = pdt, model = "random" )


summary(remethod_1)


remethod_2 <- plm(logwage ~ hours + age + female + educlevel_prehigh +educlevel_high + educlevel_further
                  + educlevel_bachelor + educlevel_masters + educlevel_phd
                  + primary + secondary + tertiary1 +tertiarycommerce
                  + overedu + underedu + otheredu 
                  + overskilled + underskilled + otherskill , data = pdt, model = "random" )

summary(remethod_2)

```



Skewness and kurtosis for logwages and wages
```{r,echo=TRUE}
# Calculate skewness and kurtosis for wages
skewness(pdt$wage)
kurtosis(pdt$wage)
# Calculate skewness and kurtosis for logwages
skewness(pdt$logwage)
kurtosis(pdt$logwage)
```

Phi-coefficients
```{r,echo=TRUE}
#Phi-coefficients 
as.data.frame(cor(data.frame(pdt$overedu, pdt$underedu, pdt$otheredu, 
               pdt$overskilled, pdt$underskilled, pdt$otherskill)))
```


